---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

pkgload::load_all()
library(knitr)
```

# coedata

<!-- badges: start -->
[![R-CMD-check](https://github.com/Church-Army/coedata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Church-Army/coedata/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/Church-Army/coedata/graph/badge.svg)](https://app.codecov.io/gh/Church-Army/coedata)
<!-- badges: end -->

coedata is a data package that serves up various datasets made available by the Church of England's [Data Services](https://www.churchofengland.org/about/data-services) team. Data sources include:

- Parish, diocese and national-level 2021 census data (Updated November 2024)
- A database of Churches, Parishes, Dioceses and other geographies (Updated November 2024)
- Parish-level data from the 2019 Index of Multiple Deprivation (Updated November 2024)

## Installation

You can install the development version of coedata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("Church-Army/coedata")
```

## Documentation

Full documentation website on: https://Church-Army.github.io/coedata

## Usage

### Using coedata to get parish-level census statistics

This is an example of how you could use coedata to get 2021 census data on general health for a selection of parishes. We'll use the neighbouring parishes of **Stifford: St Mary** and **Grays: St Peter and St Paul**.

To get this data, we'll need:

* The parish codes of those parishes
* The ONS ID of the general health dataset

We can find the ONS ID of the general health dataset by using `coe_datasets()`, which lists the ID of every census dataset that's a available through this package:

```{r coe_datasets}
coe_census_datasets() |> 
  knitr::kable()
```

We can see that general health data can be found in the dataset with ID TS037.

Luckily, we already know the parish codes for the parishes we're interested in. If you're not sure where to find the parish codes you need, please refer to [Finding parish codes].

Now that we have the ONS ID and the parish codes, we can get our census data:

```{r parish_level_data}
coe_census_parish(ons_id = "TS037", parish_codes = c(580342, 580334))
```

Those parishes are both in central London - they are very densely populated!

Note that we can also get relative statistics for the same data by setting `relative = TRUE`:

```{r parish_level_data_relative}
coe_census_parish(ons_id = "TS037", parish_codes = c(580342, 580334), relative = TRUE) 
```

For more information on getting census data, see `vignette("coe-census-data")`.

### Using coedata to get parish-level deprivation statistics

The CoE produce deprivation data for every parish in england, which can be accessed like so:

```{r deprivation_example}
coe_parish_deprivation(parish_codes = c(580342, 580334))
```

For more information on getting deprivation data, see `vignette("parish-deprivation-data")`

## Finding parish codes

When you're looking at individual parishes with `coedata`, you'll need to identify them with their unique parish codes. If you're not sure what the parish code is for a parish or church, you can either:

* Find the parish in the [interactive map](https://www.arcgis.com/home/webmap/viewer.html?webmap=67bce0ed36dd4ee0af7a16bc079aa09a) provided by the Church of England's Data Services team
* Find the church's church code by clicking 'more information' on its [A Church Near You](https://www.achurchnearyou.com) page, and then use `coe_parish_from_church()` to find it's parish code.

## Thanks and attribution

* All census data was originally provided by the [Office for National Statistics](https://www.ons.gov.uk/). ONS also maintain the [nomis](https://www.nomisweb.co.uk/) service, which was used to source some of this package's internal data.
* Parish-level census data profiles were compiled by Ken Eames, who is a senior statistical researcher at the Church of England's Data Services. Original parish-level datasets are available on the [Church of England website](https://www.churchofengland.org/about/data-services/resources-publications-and-data)
  * Ken also maintains the [parish map](https://www.arcgis.com/home/webmap/viewer.html?webmap=67bce0ed36dd4ee0af7a16bc079aa09a) from which Church data are read by this package. Thanks again, Ken!
* Thanks to [ropensci](https://github.com/ropensci) for their development and maintenance of the [nomisr](https://github.com/ropensci/nomisr) package, which was used to interface with nomis.
* Thanks to [Jenny Bryan](https://jennybryan.org/about/) at Posit for the development and maintenance of [googlesheets4](https://googlesheets4.tidyverse.org/), which provided a simple, secure way of minimising package size by storing data remotely.
