---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

pkgload::load_all()
```

# coedata

<!-- badges: start -->
[![R-CMD-check](https://github.com/Church-Army/coedata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Church-Army/coedata/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/Church-Army/coedata/graph/badge.svg)](https://app.codecov.io/gh/Church-Army/coedata)
<!-- badges: end -->

coedata is a data package that serves up various datasets made available by the Church of England's [Data Services](https://www.churchofengland.org/about/data-services) team. Data sources include:

- [x] [Parish-level census data](https://www.churchofengland.org/about/data-services/resources-publications-and-data) from ONS' 2021 Census of England and Wales (Updated November 2024)
- [x] A [database](https://services5.arcgis.com/KDRjxGRQDVgVtFTS/ArcGIS/rest/services/Churches_ACNY_Nov2024/FeatureServer) of Churches, Parishes, Dioceses and other geographies (Updated November 2024)
- [ ] ~~[Parish-level data](https://www.churchofengland.org/about/data-services/resources-publications-and-data) from the 2019 Index of Multiple Deprivation~~ (PENDING)

National-level data are sourced directly from ONS data via [nomis](https://www.nomisweb.co.uk/) (via [nomisr](https://github.com/ropensci/nomisr)).

## Installation

You can install the development version of coedata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("Church-Army/coedata")
```

## Documentation

Full documentation website on: https://Church-Army.github.io/coedata

## A simple example

Let's use `coedata` to get 2021 census data on general health for a couple of parishes. We'll use the neighbouring parishes of **Stifford: St Mary** (parish code 580342) and **Grays: St Peter and St Paul** (parish code 580334).

```{r parish_level_data}
coe_census_parish(ons_id = "TS037", parish_codes = c(580342, 580334))
```

**TS037** is the ONS ID of the dataset that contains General Health information. You can view all ONS datasets made available in this package like so:

```{r available_datasets}
coe_census_datasets()
```

Note that we can also get relative statistics for the same data by setting `relative = TRUE`:

```{r parish_level_data_relative}
coe_census_parish(ons_id = "TS037", parish_codes = c(580342, 580334), relative = TRUE)
```

## Example 2: Viewing a parish in its diocesan and national contexts

Sometimes people want to see how their parish compares to its diocese and to the nation as a whole. `coedata` contains a function that returns this data for any parish and any number of ONS datasets:

```{r parish_context_example}
coe_parish_snapshot(580342, ons_ids = "TS037")
```

## Where to find parish codes

When you're looking at individual parishes with `coedata`, you'll need to identify them with their unique parish codes. If you're not sure what the parish code is for a parish or church, you can either:

* Find the parish in the [intercative map](https://www.arcgis.com/home/webmap/viewer.html?webmap=67bce0ed36dd4ee0af7a16bc079aa09a) provided by the Church of England's Data Services team
* Find the church's church code by clicking 'more information' on its [A Church Near You](https://www.achurchnearyou.com) page, and then use `coe_parish_from_church()` to find it's parish code.

## A note on socio-economic classification labels

Most of the statistics used in this package are labelled very intuitively, using labels like `general_health_very_good` or `country_of_birth_middle_east_asia`. This is not true of the socio0economic classification data (TS062), which uses labels like `ns_sec_L1_3`. To understand these labels, please consult the table returned by `ns_sec_descriptions()`.

## Thanks and attribution

* All census data was originally rovided by the [Office for National Statistics](https://www.ons.gov.uk/). ONS also maintain the [nomis](https://www.nomisweb.co.uk/) service, which was used to source some of this package's internal data.
* Parish-level census data profiles were compiled by Ken Eames, who is a senior statistical researcher at the Church of England's Data Services. Original parish-level datasets are available on the [Church of England website](https://www.churchofengland.org/about/data-services/resources-publications-and-data)
  * Ken also maintains the [parish map](https://www.arcgis.com/home/webmap/viewer.html?webmap=67bce0ed36dd4ee0af7a16bc079aa09a) from which Church data are read by this package. Thanks again, Ken!
* Thanks to [ropensci](https://github.com/ropensci) for their development and maintenance of the [nomisr](https://github.com/ropensci/nomisr) package, which was used to interface with nomis.
* Thanks to [Jenny Bryan](https://jennybryan.org/about/) at Posit for the development and maintenance of [googlesheets4](https://googlesheets4.tidyverse.org/), which provided a simple, secure way of minimising package size by storing data remotely.

## TODO:

- [x] Replace national level data with NOMIS census data
  - Aggregating parish data causes considerable rounding error
- [x] Add data-dictionary vignettes
- [x] Create print methods for `"coe_parish_data"` class
- [x] Create methods for `dplyr` that preserve attributes in `"coe_parish_data"` objects
- [ ] Add IMD data
- [x] Ensure appropriate credit/attribution/documentation
- [x] Rename everything, perhaps including this package
