---
title: "flat_read_church_data.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

The functions on this page all help with processing data related to individual churches:

* `coe_churches()` returns a tibble of every church in the Church of England, including information on its parish, benefice, deanery, archdeaconry and diocese. Data are read directly from the [ESRI feature server](https://services5.arcgis.com/KDRjxGRQDVgVtFTS/ArcGIS/rest/services/Churches_ACNY_Nov2024/FeatureServer) kindly provided by the Church of England's Data Services team.
* `coe_esri_fields()` is a helper function that returns valid fields which can be passed to the `where` argument of `coe_churches()` to query the dataset.
* `coe_parish_from_church()` returns a parish code for each of any number of church codes. Useful for obtaining a parish code from the church codes listed at [A Church Near You](https://www.achurchnearyou.com/)

## coe_churches()

Get tidy data on Church of England churches. 

```{r function-coe_churces}
#' Get data on the Anglican geography on Church of England Churches, including their parish, benefice, arcdeaconry and diocese. 
#'
#' Reads church data from the Church of England's [Data Services ESRI feature server](https://services5.arcgis.com/KDRjxGRQDVgVtFTS/ArcGIS/rest/services/Churches_ACNY_Nov2024/FeatureServer), dropping spatial data and returning a [tibble::tibble()].
#'
#' @param where An SQL query. The default returns all churches. See details.
#' @return A tibble in which each row corresponds to a church in the Church of England
#'
#' @details
#' This function renames fields in the returned tibble, but any query passed to `where` must use the original names of the ESRI data provided by CoE data services. The names of these fields
#' can be accessed via [coe_esri_fields()]
#' 
#' @export
#' @examples
coe_churches <- function(where = "1=1") {

churches <- esri_churches(where = where, outFields = "*")

churches <- churches[c("ChurchCode", "ParLocNa", "ParishID", "BenefNa", "BenefID",
                       "DeanNa", "DeanID", "ArchdNa", "ArchdID", "DioNa", "DioID")]

names(churches) <- c("church_code", "parish_name", "parish_code", "benefice_name", "benefice_id",
                     "deanery_name", "deanery_id", "archdeaconry_name", "archdeaconry_id", "diocese_name", "diocese_number")

churches
}
```

```{r examples-coe_churches}
coe_churches(where = "ChurchCode = 608186") |>
  dplyr::glimpse()
```

```{r tests-read_church_data}
test_that("coe_churches works", {
  expect_true(inherits(coe_churches, "function"))
})
```

```{r function-esri_churches}
esri_churches <- function(where, outFields){
  req <-
  httr2::request("https://services5.arcgis.com") |> 
  httr2::req_url_path("KDRjxGRQDVgVtFTS/arcgis/rest/services/Churches_ACNY_Nov2024/FeatureServer/0/query") |> 
  httr2::req_url_query(where = where, outFields = outFields, returnGeometry = "false", f = "geojson")

  churches <- sf::st_read(req$url, quiet = TRUE)

  churches <- tibble::as_tibble(churches)
  churches <- churches[!names(churches) %in% c("geometry", "lattitude", "longitude")]
  churches
}
```

## coe_parish_from_church()

Get the parish code of any number of churches.

```{r function-coe_parish_from_church}
#' Get the parish codes of any number of Church of England churches.
#' 
#' @param church_codes A character vector of church codes
#' @return A character vector of parish codes, with the same length as `church_codes`
#' 
#' @export

coe_parish_from_church <- function(church_codes){
  
  query_codes <- paste0("(", paste(church_codes, collapse = ", "), ")")
  
  query <- paste("ChurchCode IN", query_codes)
  
  parishes <- esri_churches(where = query, outFields = "ChurchCode, ParishID")
  
  parishes$ParishID[match(church_codes, parishes$ChurchCode)]
}
```

```{r examples-coe_parish_from_church}
coe_parish_from_church(c(635041, 623001))
```

## coe_esri_fields()

This tiny function returns the names of the fields used in the Feature Server of Churches provided by CoE Data Services through ESRI.

```{r function-coe_esri_fields}
#' View fields that can be passed to the `where` argument of [coe_churches()]

#' @returns A character vector of field names
#' @export
coe_esri_fields <- function(){
  c("FID", "ChurchCode", "ChurchName", "ACNY_url", "ACNY_id", "Heritg_url",
    "LBGrade", "ParLocNa", "ParLegNa", "ParishID", "BenefNa", "BenefID", 
    "DeanNa", "DeanID", "ArchdNa", "ArchdID", "DioNa", "DioID")
}
```

```{r examples-coe_esri_fields}
coe_esri_fields()
```

