---
title: "flat_read-coe-statistics.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

The functions on this page are all used for reading 2021 ONS census data for Church of England geographies. You can browse available datasets with `coe_census_datasets()`. Once you know the ONS ID of the dataset you need, you can:

-   Get census data at a level of your choosing with the general-purpose function `coe_census()`
    -   Get parish-level data with `coe_census_parish()`
    -   Get diocese-level data with `coe_census_diocese()`
    -   Get data for all of england with `coe_census_england()`

You can also compare a parish with its diocese and the whole of england in a single table with `coe_parish_snapshot()`

## View available data with `coe_census_datasets()`

`coe_census_datasets()` displays information on all the census datasets available through this package. By default, the function returns a two-column tibble including `ons_id` and `description`, but setting `description = FALSE` will return a character vector of all valid ONS ids that can be passed to functions such as `coe_census()`.

Parish-level datasets were produced by Ken Eames, Senior Statistical Researcher at the Church of England's Data Services team. The original datasets are available on the [Church of England website](https://www.churchofengland.org/about/data-services/resources-publications-and-data). This package uses data that were last updated November 2024. If you notice that the source data have since been updated, please create a Github issue at <https://github.com/Church-Army/coedata>

```{r function-coe_census_datasets}
#' Browse all available datasets of ONS census 2021 data
#' 
#' @return A tibble of ONS ids and dataset descriptions, or a character vector of ONS ids.
#' @param description Logical. If TRUE, (the default), a tibble is returned that contains ONS ids and corresponding dataset descriptions. If false, a character vector containing only ONS ids is returned. 
#' @export

coe_census_datasets <- function(description = TRUE){
  out <- read_parish_data()
  
  out <- out[c("ons_id", "description")] |>
    unique()
  
  if(!description) out <- out[["ons_id"]]
  
  out
}

```

```{r examples-coe_census_datasets}
coe_census_datasets() |>
  dplyr::glimpse()
```

## Obtain census statistics for a parish, diocese, or the nation with `coe_census()`

Get statistics on a particular topic, for a particular geographical level, filtering as required. Levels include 'parish', 'diocese' or 'england'. Parishes can be filtered with parish codes and dioceses with diocese numbers, but national-level tables are comprised of a single row and cannot be filtered.

```{r function-coe_census}
#' Get census data for the Church of England at a parish, diocesan or national level.
#'
#' @return A [tibble::tibble()] containing census data for the Church of England  
#' @param ons_id The ONS id of required data. Must be one of [coe_census_datasets()].
#' @param level The geographical level at which data are required. Must be one of 'parish', 'diocese' or 'england'.
#' @param areas A character vector of parish codes or diocese numbers. If specified, the returned tibble will contain only data for these areas. Cannot be used with `level = 'england'`. `NA` values will be removed.
#' @param relative Logical. Should returned table contain absolute statitiscs (e.g. persons/households) or proportions for each area?
#' @export 
#'
coe_census <- function(ons_id, level, areas, relative = FALSE){
  
  ## Validate inputs
  
  ### Check ons_id is character vector with length one
  if(!is.character(ons_id) || length(ons_id) != 1) rlang::abort("Argument ons_id must be a character vector with length 1", class = "coe_bad_ons_id")
  
  ### Check ons_id's are all good
  if(!all(ons_id %in% good_ons_ids())){
    
    rlang::abort(c("Argument ons_id pertains to data that is not available through this package:", ons_id),
                 class = "coe_bad_ons_id")
  }
  
  ### level must be length 1
  if(length(level) != 1 || !is.character(level)) rlang::abort("Argument 'level' must be a character vector with length 1.",
                                                              class = "coe_bad_level")
  
  ### Check level are all good
  if(!all(level %in% good_levels())){
    good_level_txt <-
      as.character(good_levels()) |> 
      paste0("\"", foo = _, "\"")
    
    good_level_text <- paste0("c(", paste(good_level_txt, collapse = ", "), ")") 
    
    rlang::abort(c(
      paste("Argument 'level' must be one of: ", good_level_text),
      `!` = "Problematic argument:",
      `x` = paste0("level = ", level)),
      class = "coe_bad_level")
  }
  
  ## Validate areas
  ### No missing values
  if(!rlang::is_missing(areas) && any(is.na(areas))){
    rlang::warn("Ignoring missing values in argument 'areas'", class = "coe_warn_areas")
    areas <- areas[!is.na(areas)]
  }
  
  ## Get data
  out <- read_cpd_stats(ons_id, level)

  if(!rlang::is_missing(areas)){
    if(level == "parish")  out <- out[out$parish_code %in% areas, ]
    if(level == "diocese") out <- out[out$diocese_number %in% areas, ]
    if(level == "england") rlang::abort("Argument 'areas' provided but can not be used with level = 'england'",
                                        class = "coe_areas_with_england")
  }
  
  if(relative) out <- cpd_relative(out)
  
  out
}
```

```{r examples-coe_census}
coe_census(ons_id =  "TS001", level = "england")
coe_census(ons_id = "TS001", level = "diocese", areas = 1)
```

```{r tests-coe_census}
test_that("coe_census works", {
  expect_true(inherits(coe_census, "function"))
  expect_s3_class(
    coe_census("TS004", "parish"),
    c("coe_parish_data", "tbl_df")
  )
})

test_that("coe_census doesn't work when it shouldn't", {
  
  expect_error(coe_census("TS001", level = c("parish", "diocese")),
               class = "coe_bad_level",
               regexp = "length")
  
  expect_error(coe_census("TS001", level = "bad_value"),
               class = "coe_bad_level")
  
  expect_error(coe_census(c("TS001", "TS002")),
               class = "coe_bad_ons_id",
               regexp = "length")
  
  expect_error(coe_census("bad_ons_id"),
               class = "coe_bad_ons_id")
  
  expect_warning(coe_census("TS001", level = "parish", areas = NA_character_),
                 class = "coe_warn_areas")
  
  expect_error(coe_census("TS001", level = "england", areas = "yes"),
               class = "coe_areas_with_england")
})
```

## Obtain parish-level data with `coe_census_parish()`

A convenient wrapper around `coe_census()` for getting parish-level statistics for parishes of your choosing

```{r function-coe_census_parish}
#' Obtain parish-level census statistics on a given topic 
#' 
#' @param ons_id The ONS id the required topic
#' @param parish_codes Parish codes of required parishes. If empty (the default) data for all parishes are returned.
#' @param ... Other arguments passed to [coe_census()]
#' 
#'@export 
coe_census_parish <- function(ons_id, parish_codes, ...){
  
  if(!rlang::is_missing(parish_codes)) out <- coe_census(ons_id, level = "parish", areas = parish_codes, ...)
  else out <- coe_census(ons_id, level = "parish", ...)
  out
}
```

```{r examples-coe_census_parish}
coe_census_parish(ons_id = "TS062", parish_codes = "350041", relative = TRUE)
```

## Obtain diocese-level data with `coe_census_diocese()`

A convenient wrapper around `coe_census()` for getting diocese-level statistics for dioceses of your choosing

```{r function-coe_census_diocese}
#' Obtain diocese-level census statistics on a given topic
#' 
#' @param ons_id The ONS id for the required topic
#' @param diocese_numbers Diocese numbers of required dioceses. If empty (the default) data for all dioceses are returned.
#' @param ... Other arguments passed to [coe_census()]
#' 
#' @export
coe_census_diocese <- function(ons_id, diocese_numbers, ...){
  
  if(!rlang::is_missing(diocese_numbers)) out <- coe_census(ons_id, level = "diocese", areas = diocese_numbers, ...)
  else out <- coe_census(ons_id, level = "diocese", ...)
  out
}
```

```{r examples-coe_census_diocese}
coe_census_diocese(ons_id = "TS062", diocese_numbers = c(3, 7), relative = TRUE)
```

## Obtain data for all of England with `coe_census_england()`

A convenient wrapper around `coe_census()` for getting national-level stats for a topic of your choosing

```{r function-coe_census_england}
#' Obtain national level census statistics on a given topic
#' 
#' @param ons_id The ONS id for the required topic
#' @param ... Other arguments passed to [coe_census()]
#' 
#' @export
coe_census_england <- function(ons_id, ...){
  coe_census(ons_id, level = "england", ...)
}
```

```{r examples-coe_census_england}
coe_census_england(ons_id = "TS062", relative = TRUE)
```

## Compare a parish to it's diocesan and national context with `coe_parish_snapshot()`

Get a contextual 'snapshot' of any given parish for any number of census topics, comparing the parochial statistics with those of the diocese and the whole of England. A wrapper around `coe_census_parish()`, `coe_census_diocese()` and `coe_census_england()` to obtain relative statistics for each geography and combine them into a single table, iterating over census topics to return a list.

```{r function-coe_parish_snapshot}
#' Get a list of comparative table(s) of local, diocesan and national data for a given parish
#' 
#' Extracts parish, diocese and national level data fcor a given parish, then [base::rbind]s them into a single table. Iterates over `ons_ids` to output a list of such tables
#' @returns A list of `coe_parish_data` tibbles
#' @param parish_code A single parish code. Character.
#' @param ons_ids A character vector of ons_ids.
#' @param relative Logical. Should outputs be relative? Default is `TRUE`.
#' @export

coe_parish_snapshot <- function(parish_code,
                                  ons_ids = coe_census_datasets(description = FALSE),
                                  relative = TRUE){
  
  p_table <- read_parish_table()
  
  ons_ids = as.list(ons_ids)
  names(ons_ids) = ons_ids
  
  if(length(parish_code) != 1) rlang::abort("Argument 'parish_code' must have length one")
  parish_code <- as.character(parish_code)
  
  if(!parish_code %in% p_table$parish_code) rlang::abort(c("Parish code:", x = parish_code, "is not valid"))
  
  diocese_no <- p_table$diocese_number[p_table$parish_code == parish_code]
  
  par_table <- read_parish_table()
  par_table <- par_table[par_table$parish_code %in% parish_code,]
  
  parish_stats  <- lapply(ons_ids,
                          \(x){
                            stat_table <- coe_census(ons_id = x, level = "parish",  areas = parish_code, relative = relative) 
                            
                            stat_table$parish_name <- par_table$parish_name[match(stat_table$parish_code, par_table$parish_code)]
                            
                            stat_table <- coe_relocate(stat_table, c("parish_code", "parish_name"))
                            names(stat_table)[1:2] <- c("level_code", "level_name")
                            stat_table$level <- "parish"
                            stat_table <- coe_relocate(stat_table, "level")
                            stat_table
                            
                            })
  
  diocese_stats <- lapply(ons_ids,
                          \(x){
                            stat_table <- coe_census(ons_id = x, level = "diocese", areas = diocese_no, relative = relative)
                            
                            stat_table$diocese_name <- par_table$diocese_name[match(stat_table$diocese_number, par_table$diocese_number)]
                            
                            stat_table <- coe_relocate(stat_table, c("diocese_number", "diocese_name"))
                            names(stat_table)[1:2] <- c("level_code", "level_name")
                            stat_table$level <- "diocese"
                            stat_table <- coe_relocate(stat_table, "level")
                            stat_table
                            
                            })
  england_stats <- lapply(ons_ids,
                          \(x){
                            stat_table <- coe_census(ons_id = x, level = "england", relative = relative)
                            
                            stat_table$level <- "nation"
                            stat_table$level_code = NA_character_
                            stat_table$level_name = "england"
                            
                            stat_table <- coe_relocate(stat_table, c("level", "level_code", "level_name"))
                            stat_table
                            
                            })
  
  
  out <- vector("list", length(ons_ids))
  names(out) <-  names(ons_ids)
  
  for(code in names(ons_ids)){
    out[[code]] <- rbind(parish_stats[[code]], diocese_stats[[code]], england_stats[[code]])
  }
  
  out
}

```

```{r examples-coe_parish_snapshot}
coe_parish_snapshot(parish_code = "370047", ons_ids = "TS001")
```

```{r tests-coe_parish_snapshot}
test_that("coe_parish_snapshot works", {
  expect_true(inherits(coe_parish_snapshot, "function"))
  
  test <- coe_parish_snapshot("370047", ons_ids = "TS001")
  
  expect_type(test, "list")
  
  expect_equal(nrow(test[[1]]), 3)
})

test_that("stat names are consistent across levels", {
  
  census_data_copy <-
    .coedata_envir$parish_data |> 
    dplyr::select(ons_id, level, data, drive_id) |> 
    dplyr::rowwise()
  
  sheet_names <- function(data, ons_id, drive_id){
    if(is.null(data)){
      # read first row (i.e names)
      out <- 
        get_cpd_sheet(drive_id, ons_id, range = "1:1") |>
        names()
    } else out <- names(data)
    
    out
    
  }

  census_data_copy <- 
    dplyr::mutate(census_data_copy, data_names = list(sheet_names(data, ons_id, drive_id))) |> 
    tidyr::unnest(data_names) |> 
    dplyr::filter(!data_names %in% c("parish_code", "parish_name", "diocese_number", "diocese_name"))
  
  names_table <- 
    dplyr::select(census_data_copy, ons_id, level, data_names) |> 
    tidyr::pivot_wider(names_from = level, values_from = data_names, values_fn = list)
  
    dplyr::rowwise(names_table) |> 
    dplyr::mutate(
      test_diocese = list(expect_equal(parish, diocese)),
      test_england = list(expect_equal(parish, england)),
      .keep = "none"
    )
  
})
```

```{r function-coe_relocate}
coe_relocate <- function(.data, cols){
  out <- .data[,c(cols, setdiff(names(.data), cols))]
  out
}
```

## A note on socio-economic classification labels

Most of the statistics used in this package are labelled very intuitively, using labels like `general_health_very_good` or `country_of_birth_middle_east_asia`. This is not true of the socio0economic classification data (TS062), which uses labels like `ns_sec_L1_3`. To understand these labels, please consult the table returned by `ns_sec_descriptions()`.
