---
title: "flat_first.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```


```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

## Create data environment on load

```{r function-create-environment}
rlang::on_load({
  .coedata_envir <- new.env(parent = emptyenv())
  .coedata_envir$parish_data <- tibble::as_tibble(parish_data)
  .coedata_envir$parishes_2024 <- NULL
  .coedata_envir$parish_imd_2019 <- NULL
  })
```

## read_parish_data

```{r function-read_parish_data}
read_parish_data <- function(){
  .coedata_envir$parish_data
}
```

```{r tests-read_parish_data}
test_that("read_parish_data works", {
  expect(inherits(read_parish_data, "function"), "read_parish_data is not a function")
  expect_s3_class(read_parish_data(), "tbl_df")
})
```


## read_parish_table

```{r function-read_parish_table}
read_parish_table <- function(){
  p_table <- .coedata_envir$parishes_2024
  
  if(is.null(p_table)){
    p_table <- coe_churches()
    
    p_table <-
      p_table[-which(names(p_table) == "church_code")] |> 
      unique()

    .coedata_envir$parishes_2024 <- p_table
  } 
  
  p_table
}
```

## good_ons_ids

```{r function-good_ons_ids}
good_ons_ids <- function(){
  unique(read_parish_data()$ons_id)
}
```

## good_levels

```{r function-good_levels}
good_levels <- function(){
  unique(read_parish_data()$level)
}
```

## get_cpd_sheet

```{r function-get_cpd_sheet}
get_cpd_sheet <- function(drive_id, sheet){
  
  googlesheets4::gs4_deauth()
  
  googlesheets4::read_sheet(ss = drive_id, sheet = sheet)
}
```

```{r tests-get_cpd_sheet}
test_that("get_cpd_sheet works",{
  expect_true(inherits(get_cpd_sheet, "function"))
  
  will_work <- min(which(!is.na(read_parish_data()$drive_id)))

  drived <- read_parish_data()$drive_id[will_work]
  sheet  <- read_parish_data()$ons_id[will_work]

  expect_no_error(get_cpd_sheet(drived, sheet))
  })
```

## cpd_relative

```{r function-cpd_relative}
cpd_relative <- function(.data){
  if(!"coe_parish_data" %in% class(.data)) rlang::abort("'.data' must be a coe_parish_data object.")
  
  units <- attr(.data, "units")
  
  if(units == "persons")                   col <- "population"
  else if(units == "persons_sixteen_plus") col <- "population_16_and_over"
  else if(units == "households")           col <- "occupied_households"
  else rlang::abort("'units' attribute of '.data' not recognised")
  
  ignore <- c("parish_code", "parish_name", "diocese_number", "diocese_name")
  
  names_list <- names(.data)[!names(.data) %in% c(col, ignore)]
  
  for(col_name in names_list){
    .data[[col_name]] <- .data[[col_name]] / .data[[col]]
  }
  
  attr(.data, "relative") <- TRUE
  attr(.data, "units") <- paste0("Proportion of all ", attr(.data, "units"))
  .data
  
}
```

```{r tests-cpd_relative}
test_that("cpd_relative() works",{
  expect_true(inherits(cpd_relative, "function"))
  expect_no_error(cpd_relative(read_cpd_stats("TS001", "diocese")))
})
```

## read_cpd_stats

```{r function-read_cpd_stats}
#' Low level function for getting parish-level census data
#' 
#' Get 2021 ONS census data for any number of parishes, dioceses or the whole of England. Parish data is read from a Google sheet
#' 
#' @return A single tibble
#' @param ons_id Nomis code of required data set
#' @param level One of 'parish', 'diocese' or 'england'
#' 
read_cpd_stats <- function(ons_id, level){
  
  cpd_pd <- read_parish_data()

  requested_row <- which(cpd_pd$ons_id == ons_id & cpd_pd$level == level)
  stopifnot(length(requested_row) == 1)
  
  local_data <- cpd_pd$local[requested_row]
  
  if(local_data){
    out <- cpd_pd$data[requested_row][[1]]
  } else{
    # Read google sheet
    out <- get_cpd_sheet(cpd_pd$drive_id[requested_row], cpd_pd$ons_id[requested_row])
    
    ## Update data environment with sheet read from google
    .coedata_envir$parish_data[["data"]][[requested_row]] <- out
    .coedata_envir$parish_data[requested_row, "local"] <- TRUE
  }
  
    ## Assign class and units attributes
  ### Shhh! This if statement is brushing some inconsistent data formatting under the rug 
    if(!"coe_parish_data" %in% class(out)) class(out) <- c("coe_parish_data", class(out))
  
    attr(out, "ons_id") <- cpd_pd$ons_id[requested_row]
    attr(out, "description") <- cpd_pd$description[requested_row]
    attr(out, "units")    <- cpd_pd$units[requested_row]
    attr(out, "relative") <- FALSE
    
  out
}
```

```{r tests-read_cpd_stats}
test_that("read_cpd_stats works",{
  expect_true(inherits(read_cpd_stats, "function"))
  
  expect_no_error(read_cpd_stats("TS001", "diocese"))
  expect_s3_class(read_cpd_stats("TS001", "diocese"), c("coe_parish_data", "tbl_df"))
})

test_that("caching is working for parish data",{
  ## Read the data
  read_cpd_stats("TS001", "parish")

  ## Check it has cached
  tmp_dat <-  read_parish_data()
  cached <- tmp_dat$data[tmp_dat$level == "parish" & tmp_dat$ons_id == "TS001"][[1]]

  expect_s3_class(cached, c("coe_parish_data", "tbl_df"))
   })

test_that("class information is not duplicated", {
  expect_true(all(table(class(read_cpd_stats("TS001", "diocese"))) == 1))
})
```
 
## A note on socio-economic classification labels

Most of the statistics used in this package are labelled very intuitively, using labels like `general_health_very_good` or `country_of_birth_middle_east_asia`. This is not true of the socio0economic classification data (TS062), which uses labels like `ns_sec_L1_3`. To understand these labels, please consult the table returned by `ns_sec_descriptions()`.

